---
title: "Take Home Exercise 1"
author: "Zanelle Lee Wenfang"
date: "February 1, 2025"
date-modified: "last-modified"
execute:
    echo: true
    eval: true
    warning: false
    freeze: true
---

### 1. Overview

#### 1.1. Background

In this exercise, Japan's heart attack occurrences are studied alongside with several possible contributing factors. Increasing numbers of heart attacks over the years make it crucial to study the situation surrounding these occurrences, and find out if there are any correlations. Insights from this analysis will then hopefully aid as a preventive measure.

#### 1.2. Objectives

Using the dataset [Heart Attack in Japan Youth vs Adult](https://www.kaggle.com/datasets/ashaychoudhary/heart-attack-in-japan-youth-vs-adult), we attempt to understand the heart attack situation in Japan and gain insights surrounding the different age groups.

### 2. R packages

Appropriate R packages will be installed and loaded for effective data processing and transformation. The R packages used and its respective function will be displayed in the table below.

| R Package | Function |
|--------------------------|----------------------------------------------|
| [tidyverse](http://tidyverse.org/packages/) | A collection of packages to do various types data analyses. |
| [SmartEDA](https://cran.r-project.org/web/packages/SmartEDA/vignettes/SmartEDA.html) | For initial exploratory analysis describing the structure and relationships present in the data without lengthy R codes. |
| [ggdist](https://mjskay.github.io/ggdist/reference/ggdist-package.html) | For visualizations of distributions and uncertainty. |
| ggpubr |  |
|  |  |
|  |  |

```{r}
pacman::p_load(tidyverse, SmartEDA, ggdist, ggpubr)
```

### 3. Data

#### 3.1. Loading the data

The dataset is first loaded using *read_csv*.

```{r}
japan_data <- read_csv("japan_heart_attack_dataset.csv")
```

#### 3.2. Data overview and selection

First, we look at the dataset to see if there are any missing or duplicated values. Using *ExpData*, there are 30,000 rows of data with no missing values nor duplicated rows. It should also be noted that there are 15 extra columns of numerical values that are meaningless as they seem to be the results of certain calculations which are not stated.

::: panel-tabset
## Summary Table 1

```{r}
#| echo: false 
stats1 <- ExpData(data=japan_data, type=2)
stats1
```

## Code

```{r}
#| eval: false 
stats1 <- ExpData(data=japan_data, type=2)
stats1
```
:::

::: panel-tabset
## Summary Table 2

```{r}
#| echo: false
stats2 <- ExpData(data=japan_data, type=2)
stats2
```

## Code

```{r}
#| eval: false
stats2 <- ExpData(data=japan_data, type=2)
stats2
```
:::

#### 3.3. Data preparation

##### 3.3.1. Selecting desired columns

As there is no information on these columns of data, these will be filtered away from the dataset and leaving 17 columns in the new filtered dataset, *filtered_data*.

```{r}
filtered_data <- select(japan_data, Age, Gender, Region, Smoking_History
, Diabetes_History, Hypertension_History, Cholesterol_Level, Physical_Activity, Diet_Quality,	Alcohol_Consumption,	Stress_Levels,	BMI,	Heart_Rate,	Systolic_BP,	Diastolic_BP,	Family_History,	Heart_Attack_Occurrence)
```

##### 3.3.2. Adjusting variables

###### 3.3.2.1. Converting variables to suitable forms

```{r}
filtered_data <- filtered_data %>% 
  mutate(Physical_Activity=(recode(Physical_Activity, "Low"="1", "Moderate"="2", "High"="3")))
```

```{r}
filtered_data <- filtered_data %>% 
  mutate(Diet_Quality=(recode(Diet_Quality, "Poor"="1", "Average"="2", "Good"="3")))
```

```{r}
filtered_data <- filtered_data %>% 
  mutate(Alcohol_Consumption=(recode(Alcohol_Consumption, "None"="0", "Low"="1", "Moderate"="2", "High"="3")))
```

###### 3.3.2.2. Changing variable data types

Looking at the character rows, it is seen that they have fixed input and should be converted to factors using *mutate()* to reduce errors and improve efficiency. To check that this is done correctly, *glimpse()* is used.

```{r}
cols <- c("Gender", "Region", "Physical_Activity", "Diet_Quality", "Alcohol_Consumption", "Smoking_History"
, "Diabetes_History", "Hypertension_History", "Family_History",	"Heart_Attack_Occurrence")
filtered_data <- filtered_data %>%
  mutate(across(cols, as.factor))

glimpse(filtered_data)
```

##### 3.3.3. Converting variables to suitable forms

###### 3.3.3.1. Binning *Age* into Age Groups

To be able to gain better insights from the dataset, it was decided that *Age* should be binned into 4 separate bins - Youth, Adult, Middle-aged adult and Senior Citizen. The differences between the various age groups might be more visible than if it was kept as a continuous variable.

::: {.callout-tip title="Age Classification"}
For this analysis, the age classification would be as such:

|                   |                   |
|-------------------|-------------------|
| Youth             | \<= 25 years old  |
| Adult             | 26 - 44 years old |
| Middle-aged adult | 45 - 49 years old |
| Senior Citizen    | \>= 60 years old  |
:::

```{r}
binned_data <- filtered_data %>% mutate(Age_Group = cut(Age, breaks=c(0, 24, 44, 59, 80)))

binned_data <- binned_data %>% 
  relocate(Age_Group, .after = Age)

binned_data <- binned_data %>% 
  mutate(Age_Group=(recode(Age_Group, "(0,24]"="Youth", "(24,44]"="Adult", "(44,59]"="Middle-aged adult", "(59,80]"="Senior Citizen")))
glimpse(binned_data)
```

::: callout-caution
It should be noted that several columns have measurements that consist of seemingly impossible values.

The *BMI* column consists of single digit values. However, the [World Health Organisation](https://apps.who.int/nutrition/landscape/help.aspx?menu=0&helpid=420#:~:text=BMI%20%3C17.0%3A%20moderate%20and%20severe,BMI%20%E2%89%A525.0%3A%20overweight) states that BMI values of \<17.0 indicates moderate and severe thinness in adult populations and BMI \<16.0 is known to be associated with a markedly increased risk for ill-health, poor physical performance, lethargy and even death; thus making this cut-off point to be a valid extreme limit.

The [minimum systolic blood pressure and diastolic blood pressure](https://www.ncbi.nlm.nih.gov/books/NBK268/#:~:text=The%20minimum%20acceptable%20blood%20pressure,be%20great%20variation%20between%20patients.) is also stated to be 90 mm Hg systolic and 60 mm Hg diastolic while the lowest values in the dataset go as low as 56.2 mm Hg and 40.0 mm Hg (3s.f.) respectively.

However, as the dataset did not clarify the units used as well as calculation methods to obtain these values. Thus, these values are not removed from the dataset but this discrepency should be taken into account.
:::

### 4. Data exploration and analysis

#### 4.1. Exploratory Data Analysis

##### 4.1.1. Grouping similar variables

For more efficient analysis of the dataset, it is decided that the variables should be split to **Uncontrollable**, **Lifestyle** and **Measurements**. The way that it is split is as per follows:

| Uncontrollable | Lifestyle           | Measurements         |
|----------------|---------------------|----------------------|
| Gender         | Smoking_History     | Diabetes_History     |
| Region         | Physical_Activity   | Hypertension_History |
| Family_History | Diet_Quality        | Cholesterol_Level    |
|                | Alcohol_Consumption | BMI                  |
|                | Stress_Levels       | Heart_Rate           |
|                |                     | Systolic_BP          |
|                |                     | Diastolic_BP         |

*Age* and *Heart_Attack_Occurrences* are not included in the table above as they are considered the target variables in this data. However, if they were to be placed, they would fall under **Uncontrollable**.

##### 4.1.2. Age vs Heart Attack Occurrence

::: {.callout-note collapse="true" title="See codes"}
```{r}
#| eval: false
ggplot(data=binned_data, aes(x=Age_Group, fill=Heart_Attack_Occurrence)) +
  geom_bar(width=.8, alpha = 0.7, position="dodge") +
  ggtitle(" ") + xlab("Age Group") + labs(fill="Heart Attack Occurrence")
```
:::

```{r}
#| echo: false
ggplot(data=binned_data, aes(x=Age_Group, fill=Heart_Attack_Occurrence)) +
  geom_bar(width=.8, alpha = 0.7, position="dodge") +
  ggtitle(" ") + xlab("Age Group") + labs(fill="Heart Attack Occurrence")
```

```{r}
#| echo: false
Y_HA <- nrow(binned_data[binned_data$Age_Group=="Youth" & binned_data$Heart_Attack_Occurrence=="Yes",])/ nrow(binned_data[binned_data$Age_Group=="Youth",]) *100

A_HA <- nrow(binned_data[binned_data$Age_Group=="Adult" & binned_data$Heart_Attack_Occurrence=="Yes",])/ nrow(binned_data[binned_data$Age_Group=="Adult",]) *100

M_HA <- nrow(binned_data[binned_data$Age_Group=="Middle-aged adult" & binned_data$Heart_Attack_Occurrence=="Yes",])/ nrow(binned_data[binned_data$Age_Group=="Middle-aged adult",]) *100

S_HA <- nrow(binned_data[binned_data$Age_Group=="Senior Citizen" & binned_data$Heart_Attack_Occurrence=="Yes",])/ nrow(binned_data[binned_data$Age_Group=="Senior Citizen",]) *100
```

The percentages of heart attack occurrences in each age group can also be seen below.

```{r}
#| echo: false
percentage <- c("Youth"=Y_HA, "Adult"=A_HA, "Middle-age adult"=M_HA, "Senior Citizen"=S_HA)
percentage

```

##### 4.1.3. Uncontrollable VS Heart Attack Occurrence

::: {.callout-note collapse="true" title="See codes"}
```{r}
#| eval: false
p1 <- ggplot(data=binned_data, aes(y=Heart_Attack_Occurrence, x=Age, fill=Gender)) +
  geom_violin(alpha=0.3, position="identity") +
  ylab("Heart Attack Occurrence") + xlab("Age")

p2 <- ggplot(data=binned_data, aes(y=Heart_Attack_Occurrence, x=Age, fill=Region)) +
  geom_violin(alpha=0.3, position="identity") +
  ylab("Heart Attack Occurrence") + xlab("Age")

p3 <- ggplot(data=binned_data, aes(y=Heart_Attack_Occurrence, x=Age, fill=Family_History)) +
  geom_violin(alpha=0.3, position="identity") +
  ylab("Heart Attack Occurrence") + xlab("Age") + labs(fill="Family History")

figure <- ggarrange(p1 + rremove("ylab") + rremove("xlab"), p2 + rremove("xlab"), p3 + rremove("ylab"),
                    ncol = 1, nrow = 3)
figure
```
:::

```{r}
#| echo: false
p1 <- ggplot(data=binned_data, aes(y=Heart_Attack_Occurrence, x=Age, fill=Gender)) +
  geom_violin(alpha=0.3, position="identity") +
  ylab("Heart Attack Occurrence") + xlab("Age")

p2 <- ggplot(data=binned_data, aes(y=Heart_Attack_Occurrence, x=Age, fill=Region)) +
  geom_violin(alpha=0.3, position="identity") +
  ylab("Heart Attack Occurrence") + xlab("Age")

p3 <- ggplot(data=binned_data, aes(y=Heart_Attack_Occurrence, x=Age, fill=Family_History)) +
  geom_violin(alpha=0.3, position="identity") +
  ylab("Heart Attack Occurrence") + xlab("Age") + labs(fill="Family History")

figure <- ggarrange(p1 + rremove("ylab") + rremove("xlab"), p2 + rremove("xlab"), p3 + rremove("ylab"),
                    ncol = 1, nrow = 3)
figure
```

##### 4.1.4. Lifestyle VS Heart Attack Occurrence

::: panel-tabset
## Smoking

```{r}
ggplot(data=binned_data, aes(x=Age_Group, color=Smoking_History, fill=Heart_Attack_Occurrence)) +
  geom_bar(alpha = 0.2, width= 0.8, stat="count", position="fill") +
  scale_y_continuous(labels = scales::percent) + xlab("Age Group") + 
  ylab("Percentage with Smoking History") + labs(color="Smoking_History", fill="Heart Attack Occurrences")
```

## Physical_Activity

```{r}
ggplot(data=binned_data, aes(x=Age_Group, color=Physical_Activity, fill=Heart_Attack_Occurrence)) +
  geom_bar(alpha = 0.2, width= 0.8, stat="count", position="fill") +
  scale_y_continuous(labels = scales::percent) + xlab("Age Group") + 
  ylab("Percentage varying levels of Physical_Activity") + labs(color="Physical_Activity", fill="Heart Attack Occurrences")
```

## Diet_Quality

```{r}
ggplot(data=binned_data, aes(x=Age_Group, color=Diet_Quality, fill=Heart_Attack_Occurrence)) +
  geom_bar(alpha = 0.2, width= 0.8, stat="count", position="fill") +
  scale_y_continuous(labels = scales::percent) + xlab("Age Group") + 
  ylab("Percentage of varying Diet Quality") + labs(color="Diet Quality", fill="Heart Attack Occurrences")
```

## Alcohol_Consumption

```{r}
ggplot(data=binned_data, aes(x=Age_Group, color=Alcohol_Consumption, fill=Heart_Attack_Occurrence)) +
  geom_bar(alpha = 0.2, width= 0.8, stat="count", position="fill") +
  scale_y_continuous(labels = scales::percent) + xlab("Age Group") + 
  ylab("Percentage of varying Alcohol_Consumption") + labs(color="Alcohol_Consumption", fill="Heart Attack Occurrences")
```
:::

##### 4.1.5. Measurements VS Heart Attack Occurrence
