---
title: "Take Home Exercise 1"
author: "Zanelle Lee Wenfang"
date: "February 1, 2025"
date-modified: "last-modified"
execute:
    echo: true
    eval: true
    warning: false
    freeze: true
---

### 1. Overview

In this exercise,

### 2. R packages

Appropriate R packages will be installed and loaded for effective data processing and transformation. The R packages used and its respective function will be displayed in the table below.

| R Package | Function |
|----|----|
| tidyverse | A collection of packages to do various types data analyses. |
| [SmartEDA](https://cran.r-project.org/web/packages/SmartEDA/vignettes/SmartEDA.html) | For initial exploratory analysis describing the structure and relationships present in the data without lengthy R codes. |
| knitr |  |
| ggthemes |  |
|  |  |
|  |  |

```{r}
pacman::p_load(tidyverse, SmartEDA)
```

### 3. Data

#### 3.1. Loading the data

Using the dataset [Heart Attack in Japan Youth vs Adult](https://www.kaggle.com/datasets/ashaychoudhary/heart-attack-in-japan-youth-vs-adult), we attempt to understand the heart attack situation in Japan and gain insights surrounding the different age groups .

```{r}
japan_data <- read_csv("japan_heart_attack_dataset.csv")
```

#### 3.2. Data overview

::: callout-note
For this analysis, the age classification would be as such:

Youth \<=25 years old

Adult 26-44 years old

Middle aged adults 45-59 years old

Senior citizens \>= 60 years old
:::

First, we look at the dataset to see if there are any missing or duplicated values. Using *ExpData*, there are 30,000 rows of data with no missing values nor duplicated rows.

```{r}
stats1 <- ExpData(data=japan_data, type=1)
stats2 <- ExpData(data=japan_data, type=2)
```

### 5. Task
